
\input{header}

\begin{document}


\title{Lecture 13}

\date{May 20, 2024}

\maketitle

	In this lecture, we prove the first part of the localization theorem. Throughout this lecture, we write $X:=\Fl_G$.

\section{Fibers of the localization functor}

	In this section, we prove the following result.

		\begin{prop}
		\label{prop-fiber-localization}
		For $M \in U(\mfg)\hmod$ and any closed point $x\in X$, we have
		\[
			\Loc(M)|_x \simeq M_{\stab_\mfg(x)},
		\]
		where $\stab_\mfg(x)$ is the stablizer of $\mfg$ at $x$, i.e., $\stab_\mfg(x):=\ker( \mfg \to \mCT(X) \to \mCT_{X,x}  )$.
	\end{prop}

	\begin{rem}
		It is easy to see $\stab_\mfg(x)$ is the Borel subalgebra of $\mfg$ corresponding to the closed point $x\in X$ (see [Lecture 12, Construction 1.7]).
	\end{rem}

	\proof
		By definition,
		\[
			\Loc(M)|_x \simeq \Gamma(X,k_x \ot_{\mCO_{X}} \mCD_{X} \ot_{\ul{U(\mfg)}} \ul{M}) \simeq \Gamma(X, \delta_x \ot_{ \ul{U(\mfg)} } \ul{M}) \simeq \Gamma(X,\delta_x) \ot_{ U(\mfg) } M,
		\]
		where $\delta_x$ is the Delta right $\mCD$-module in [Lecture 11, Exercise 6.8]. By \emph{loc.cit.}, $\delta_x$ has a unique global section $\Dirac_x\in  \Gamma(X,\delta_x)$ such that $\Dirac_x \cdot f = f(x)\Dirac_x$ for any local section $f$ of $\mCO_{X}$. It follows for any vector field $\partial$ with $\partial|_x =0$, we have $\Dirac_x \cdot \partial = 0$ because locally we can write $\partial=\sum f_k \partial_k$ with $f_k(x)=0$. In particular, the right $U(\mfg)$-action on $\Dirac_x$ annihilates $\stab_\mfg(x) \subset \mfg \subset U(\mfg)$. In other words, we have a right $U(\mfg)$-linear map
		\[
			k\ot_{U(\stab_\mfg(x))} U(\mfg) \to \Gamma(X, \delta_x ),\; 1\ot u \mapsto \Dirac_x \cdot u. 
		\]
		It is easy to see both sides have natural filtrations induced respectively by the PBW filtrations on $U(\mfg)$ and $\mCD_{X}$, and the above map is compatible with the filtrations. Taking associated graded spaces, we only need to show the following obtained map is an isomorphism
		\[
			\Sym^\bullet( \mfg/\stab_\mfg(x) ) \to \Sym^\bullet(\mCT_{X,x}).
		\]
		Unwinding the definitions, this map is induced by the isomorphism $\mfg/\stab_\mfg(x) \simeq \mCT_{X,x}$.

	\qed

	\begin{rem}
		As can be seen from the proof, Proposition \ref{prop-fiber-localization} remains true if $X$ is replaced by any homogenous space under $G$.
	\end{rem}

	\begin{rem}
		As can be seen from the proof, Proposition \ref{prop-fiber-localization} remains true for derived categories and derived functors. In other words, the derived fiber of $\Loc(M)$ at $x$ can be identified with the derived coinvariance of $M$ for $\stab_\mfg(x)$.

	\end{rem}

	Let $e\in X \simeq G/B$ be the closed point corresponding to the chosen Borel subgroup $B$. In the above proof, we have shown
	\[
		 \Gamma(X, \delta_e ) \simeq k\ot_{U(\mfb)} U(\mfg)
	\]
	as right $U(\mfg)$-modules. Note that the RHS is the ``\emph{right} Verma module'' with highest weight $0$. As stated in the localization theorem, we can produce the (left) Verma module $M_{-2\rho}$ with highest weight $-2\rho$ if using the left $\mCD$-module corresponding to $\delta_e$. The following exercise gives a direct proof to this fact.

	\begin{exe}
		\label{exe-delta-Verma-2rho}
		This is \red{Homework 6, Problem 4}. 
		In above, let $\delta_e^l\simeq \delta_e\ot \omega_{X}^{-1}$ be the left $\mCD$-module corresponding to $\delta_e$. Consider the left $U(\mfg)$-module $V:=\Gamma(X, \delta_e^l)$.
		
		\begin{itemize}
			\item[(1)]
				Prove: there is a \emph{canonical} isomorphism
				\[
					\delta_e^l \simeq \mCD_{X} \ot_{\mCO_{X}} \ell,
				\]
				where $\ell$ is the fiber of $\omega_{X}^{-1}$ at $e$, viewed as a skyscrapter sheaf.
			\item[(2)]
				Let $ \ell \inj V$ be the injection induced by taking global sections for the embedding $\mCO_{X} \ot_{\mCO_{X}} \ell \inj \mCD_{X} \ot_{\mCO_{X}}\ell$. Prove: this line in $V$ is a weight subspace of weight $-2\rho$\footnote{Hint: $\ell \simeq \wedge^d \mCT_{X,e} $ and $\mCT_{X,e}\simeq \mfn^-$.}.
			\item[(3)]
				Prove: the subalgebra $\mfb \subset \mfg$ stablizes the line $\ell\subset V$\footnote{Hint: consider the PBW filtration of $\mCD_X$ and the induced filtration on $V$. Show that $\mfb \ot \ell \to V$ factors through $\mbF^{\le 1} V$ and the composition $b\ot \ell \to \mbF^{\le 1} V \to \gr^1 V$ is zero.}.
			\item[(4)]
				Construct a $U(\mfg)$-linear map 
				\[
					M_{-2\rho} \to V
				\]
				and prove it is an isomorphism.
		\end{itemize}

	\end{exe}



\section{The ring \texorpdfstring{$\mCD(X)$}{D(Fl)}}
	

	\begin{prop}
		The homomorphism $a:U(\mfg) \to \mCD(X)$ factors through $U(\mfg)_{\chi_0}$.
	\end{prop}

	\proof
		We only need to show $a(z)=0$ for any $z\in \ker(\chi_0)\subset Z(\mfg)$. We only need to show for any closed point $x\in X$, the composition
		\[
			\ker(\chi_0) \to \mCD(X) \to \Gamma(X, k_x\ot_{\mCO_X} \mCD_X)
		\]
		is zero. By the proof of Proposition \ref{prop-fiber-localization}, this map can be identified with
		\[
			\ker(\chi_0) \to k\ot_{U(\mfb_x)} U(\mfg),
		\]
		where $\mfb_x = \stab_\mfg(x)$ is the Borel subalgebra corresponding to $x$. Note that the ideal $\ker(\chi_0)\subset Z(\mfg)$ does not depond on the choice of any Borel subalgebra: it is the character for the trivial representation. Hence we only need to show $\ker(\chi_0) \to k\ot_{U(\mfb^-)} U(\mfg)$ is the zero map. But this follows from the Harish-Chandra embedding
		\[
			Z(\mfg) \to U(\mfg) \to k\ot_{U(\mfn^-)} U(\mfg) \ot_{U(\mfn)} k\simeq U(\mft).
		\]

	\qed

	\begin{rem}
		Alternatively, we can use left $\mCD$-modules and reduce to show 
		\[
			\ker(\chi_0) \to \mCD(X) \to \Gamma(X,  \mCD_X \ot_{\mCO_X} k_e)
		\]
		is zero. By Exercise \ref{exe-delta-Verma-2rho}, the RHS is non-canonically isomorphic to $M_{-2\rho}$\footnote{Such an isomorphism depends on a trivialization of the line $\ell$, i.e., a choice of vector in it.} and the above map can be identified with the action map on a highest weight vector. Then the claim follows from $\varpi(-2\rho) = \chi_0$.
	\end{rem}

	To prove the obtained homomorphism
	\[
		U(\mfg)_{\chi_0} \to \mCD(X)
	\]
	is an isomorphism, we consider filtrations on both sides.

	\begin{constr}
		The PBW filtration on $U(\mfg)$ induces a filtration on $U(\mfg)_{\chi_0}$. The surjection $U(\mfg) \to U(\mfg)_{\chi_0}$ induces a surjection $\Sym^\bullet(\mfg) \to \gr^\bullet (U(\mfg)_{\chi_0})$ which sends $\ker(\gr^\bullet(Z(\mfg))\to k)$ to $0$. Recall $\gr^\bullet (Z(\mfg)) \simeq \Sym^\bullet(\mfg)^\mfg$ ([Lecture 5, Lemma 3.2]). Hence we obtain a surjection
		\[
			\mCO( \mfg^*\mt_{\mfg^*/\!/G} 0 ) \simeq \Sym^\bullet(\mfg) \ot_{  \Sym^\bullet(\mfg)^\mfg } k \surj \gr^\bullet (U(\mfg)_{\chi_0}).
		\] 
		Note that \emph{a priori} we do not know this is an isomorphism.

	\end{constr}
	
	\begin{constr}
		On the other hand, the short exact sequences $0 \to \mbF^{\le k-1} \mCD_X \to  \mbF^{\le k} \mCD_X \to \Sym_{\mCO_X}^k \mCT_X \to 0$ induce 
		\[
			0 \to \Gamma(X, \mbF^{\le k-1} \mCD_X) \to \Gamma(X, \mbF^{\le k} \mCD_X) \to \Gamma(X,  \Sym_{\mCO_X}^k \mCT_X )
		\]
		and therefore an injection
		\[
			\gr^\bullet \mCD(X) \inj \Gamma(X,  \Sym_{\mCO_X}^\bullet \mCT_X ) \simeq \mCO(T^*X),
		\]
		where $T^*X \simeq \Spec_X( \Sym_{\mCO_X}^\bullet \mCT_X   )$ is the cotangent bundle on $X$. Note that \emph{a priori} we do not know this is an isomorphism.

	\end{constr}

	
	Combining the above constructions, we obtain homomorphisms
	\[
		\mCO( \mfg^*\mt_{\mfg^*/\!/G} 0 ) \surj \gr^\bullet (U(\mfg)_{\chi_0}) \to \gr^\bullet \mCD(X) \inj\mCO(T^*X).
	\]
	We only need to show this composition is an isomorphism. This composition corresponds to a map
	\begin{equation}
		\label{eqn-Springer-resolution}
		T^*X \to  \mfg^*\mt_{\mfg^*/\!/G} 0
	\end{equation}
	which will be studied in the next section.

	\begin{rem}
		The map $T^*X\to \mfg^*$, which is the (algebro-geometric) dual of $\mfg \to \mCT(X)$ is called the \textbf{moment map}.

	\end{rem}

\section{Nilpotent cone and the Springer resolution}
	
	In this and the next sections, we study the map \eqref{eqn-Springer-resolution}. I recommend \cite[Section 3]{CG} for these contents.


	Recall we have an identification $\mfg \simeq \mfg^*$ provided by the Killing form. Also recall $\mfg^*/\!/G\simeq \mfg/\!/G \simeq \mft/\!/W$ are isomorphic to an affine space of dimension equal to $\dim(\mft)$ (see [Lecture 6]).

	We first describe the target of \eqref{eqn-Springer-resolution}.

	\begin{defn}
		Define $\mCN$ to be the fiber product
		\[
			\xymatrix{
				\mCN \ar[r] \ar[d] & \mfg \ar[d] \\
				0 \ar[r] & \mfg/\!/G,
			}
		\]
		and call it the \textbf{nilpotent cone} of $\mfg$.
	\end{defn}

	\begin{rem}
		\label{rem-Nilp-CM}
		By Kostant's theorem ([Lecture 6, Corollary 1.15]), the projection map $\mfg \to \mfg/\!/G$ is flat. Recall regular immersions are closed under flat base-changes. Hence $\mCN \to \mfg$ is a regular immersion. In particular, $\mCN$ is Cohen--Macaulay.
	\end{rem}

	\begin{rem}
		We have $\dim(\mCN) = \dim(\mfg)-\dim(\mft)$.
	\end{rem}

	\begin{warn}
		The nilpotent cone $\mCN$ is always singular.
	\end{warn}

	The name ``nilpotent cone'' is justified by the following result:

	\begin{prop}
		A closed point of $\mfg$ is contained in $\mCN$ iff it is an nilpotent element.
	\end{prop}

	\proof
		Recall for any Borel pair $(\mfb,\mft)$, we have a commutative diagram (see [Lecture 5, (4.2)])
		\[
			\xymatrix{
				\mfb \ar[r] \ar[d] & \mfg/\!/G \\
				\mft \ar[r] & \mft/\!/W. \ar[u]_-\simeq
			}
		\]
		Also, $W$ acts transitively on the fibers of the map $\mft\to \mft/\!/W$ at the closed points ([Lecture 6, Proposition 1.1]). It follows that a closed point $v\in \mfb$ is sent to $0\in \mfg/\!/G$ iff it is sent to $0\in \mft$. The latter condition is equivalent to $v$ being nilpotent. Now the claim follows from the fact that any element of $\mfg$ is contained in some Borel subalgebra.

	\qed

	\begin{rem}
		We will see $\mCN$ is reduced (and even normal) and therefore it can be characterized by the above proposition.
	\end{rem}

	Now we describe the source of \eqref{eqn-Springer-resolution}. Note that $T^*X$ is smooth because $X$ is so.

	\begin{prop}
		Consider the obvious projection $T^* X \to X$ and the moment map $T^*X \to \mfg^*$. The obtained map
		\begin{equation}
			\label{eqn-cotagent-as-closed-sub}
			T^*X \to X\mt \mfg^*
		\end{equation}
		is a closed embedding. Moreover, via the identification $\mfg\simeq \mfg^*$, a closed points $(x,v)\in X\mt \mfg$ is contained in $T^*X$ iff $v\in \mfn_x:=[\mfb_x,\mfb_x]$, where $\mfb_x$ is the Borel subalgebra corresponding to $x$.
	\end{prop}

	\proof
		Let $x\in X$ be a closed point. We have a ``realizing'' map $X\simeq G/B_x$, $x\mapsto B_x/B_x$. It follows that $\mCT_{X,x}\simeq \mfg/\mfb_x $ and therefore $\mCT^*_{X,x}\simeq (\mfg/\mfb_x)^* $. By definition, the fiber of \eqref{eqn-cotagent-as-closed-sub} at $x\in X$ is the obvious map $(\mfg/\mfb_x )^* \to \mfg^*$ which is a closed embedding.

		In general, a linear map between two vector bundles on $X$ is a closed embedding iff its fiber at any closed point $x\in X$ is a closed embedding. Therefore \eqref{eqn-cotagent-as-closed-sub} is a closed embedding.

		Now the second claim follows from the isomorphism $(\mfg/\mfb_x)^*\simeq \mfn_x$.

	\qed

	\begin{rem}
		\label{rem-Springer-coordinate}
		One can find local trivialization of the vector bundle $T^*X \to \Fl_G$ as follows. Let $x,x^- \in X$ be closed points such that the Borel subalgebras $\mfb_x$ and $\mfb_{x^-}$ intersect transversally. Using the Bruhat decomposition, the $(N_{x^-})$-orbit of $x$ is open, and we denote it by $U_{x,x^-}$. It follows that the commposition
		\[
			T^* X \to X\mt \mfg^* \to X \mt \mfn_{x^-}^* \simeq X \mt \mfn_x
		\]
		is an isomorphism when restricted to the open subscheme $U_{x,x^-}\subset X$. Indeed, this follows fromm the fact that for any $y\in U_{x,x^-}$, the composition $\mfb_{x^-} \to \mfg \to \mfg/\mfb_y$ is an isomorphism.
	\end{rem}

	\begin{defn}
		We write $\wt{\mCN}:=T^* X $ can call the map \eqref{eqn-Springer-resolution}
		\[
			\mfp: \wt\mCN \to \mCN.
		\]	
		the \textbf{Springer resolution of the nilpotent cone}.
	\end{defn}

	\begin{lem}
		\label{lem-Springer-proper-surj}
		The map $\mfp:\wt\mCN \to \mCN$ is proper and surjective.
	\end{lem}

	\proof
		We have a commutative diagram
		\[
			\xymatrix{
				\wt\mCN \ar[r]\ar[d] & X\mt \mfg \ar[d] \\
				\mCN \ar[r] & \mfg.
			}
		\]
		The top horizontal map is proper because it is a closed embedding. The right vertical map is proper because $X$ is complete. Hence the composition $\wt\mCN\to \mfg$ is proper. Since $\mCN \to \mfg$ is separated, the map $\wt\mCN \to \mCN$ is also proper.

		It remains to show $\mfp$ is surjective on closed points. This follows from the fact that any (nilpotent) element in $\mfg$ is contained in some Borel subalgebras.

	\qed

	\begin{cor}
		\label{cor-Nilp-irre}
		The scheme $\mCN$ is irreducible.
	\end{cor}

	We will see $\mfp: \wt\mCN \to \mCN$ is a resolution of singularities. For example, in the case of $\SL_2$, we have:

	\begin{exe}
		This is \red{Homework 6, Problem 5}. For $G=\SL_2$, prove $\mfp: \wt\mCN \to \mCN$ is the blow-up of $\mCN$ at the point $0\in \mCN$.
	\end{exe}

	\begin{rem}
		The Springer resolution plays a central role in geometric representation theory.
	\end{rem}

\section{Kostant's theorem}

	Our goal is to prove the following result.

	\begin{thm}[Kostant]
		\label{thm-Kostant-Springer}
		The map $\wt\mCN \to \mCN$ induces an isomorphism $\mCO(\mCN) \xrightarrow{\sim} \mCO(\wt\mCN)$.
	\end{thm}

	\begin{rem}
		In fact, one can show the \emph{derived} direct image functor $\mfp_*:D(\mCO_{\wt\mCN}\hmod_\qc) \to D(\mCO_{\mCN}\hmod_\qc)$ sends $\mCO_{\wt\mCN}$ to $\mCO_{\mCN}$. The proof of this stronger result is an elaboration of the proof of Theorem \ref{thm-Kostant-Springer} displayed below, with the help of the (derived non-flat) base-change isomorphisms. See \cite[Section 7]{G} for more details.

	\end{rem}

	Note that the above theorem implies the first part of the localization theorem.

	\begin{cor}
		\label{cor-U-vs-D}
		The homomorphism $U(\mfg)_{\chi_0} \to \mCD(X)$ is an isomorphism.
	\end{cor}

	\proof
		By the discussion in previous sections, we only need to show $\mCO(\mCN) \to \mCO(\wt\mCN)$ is an isomorphism, which is Kostant's theorem.

	\qed

	To prove Kostant's theorem, we need more geometric inputs.

	\begin{propdefn}
		\label{propdefn-Groth-alter}
		There is a unique reduced closed subscheme $\wt\mfg$ of $X\mt \mfg$, called the \textbf{Grothendieck's alteration}, whose closed points are those $(x,v)$ satisfying $v\in \mfb_x$. Moreover, $\wt\mfg$ is smooth.
	\end{propdefn}

	\proof[Sketch]
		It is easy to show $v\in \mfb_x$ is a closed condition and therefore defines a reduced closed subscheme $\wt\mfg$. Also, as in Remark \ref{rem-Springer-coordinate}, the commposition 
		\[
			\wt{\mfg} \to X\mt \mfg \to X \mt \mfg/\mfn_{x^-} \simeq X \mt \mfb_x
		\] 
		is an isomorphism when restricted to the open subscheme $U_{x,x^-}\subset X$. This implies $\wt{\mfg}$ is smooth.

	\qed



	\begin{lem}
		\label{lem-Groth-alter-Cartan}
		There exists a Cartesian square
		\[
			\xymatrix{
				\wt{\mCN} \ar[r] \ar[d] & \wt{\mfg} \ar[d] \\
				0 \ar[r] & \mft^\abs,
			}
		\]
		where $\mft^\abs$ is the \emph{abstact} Cartan Lie algebra (see Appendix \ref{app-abs}). Moreover, the vertical maps are smooth.

	\end{lem}

	\proof[Sketch]
		We have an obvious injective map $\wt{\mCN} \to \wt{\mfg}$ between vector bundles on $X$. By Remark \ref{rem-Springer-coordinate} and the proof of Proposition-Definition \ref{propdefn-Groth-alter}, this map can be identified with $X\mt \mfn_x \to X\mt \mfb_x $ when restricted to the open subscheme $U_{x,x'} \subset X$. Hence the quotient bundle can be identified with $X \mt \mft_x \simeq X\mt \mft^\abs$ over $U_{x,x'}$, where we used the realizing isomorphism $\mft^\abs \to \mft_x$.

		One can show these identifications can be glued into a short exact sequence of vector bundles over $X$:
		\[
			0 \to \wt\mCN \to \wt\mfg \to X \mt \mft^\abs \to 0,
		\]
		which makes the desired claim manifest.

	\qed

	\begin{notn}
		Let $\mfg_\rss \subset \mfg_\reg \subset \mfg$ be the open subschemes whose closed points are regular semisimple (resp. regular\footnote{Recall an element $v\in \mfg$ is regular if its centralizer is of minimal dimension, which is $\dim(\mft^\abs)$.}) elements in $\mfg$. Let $\wt{\mfg}_\rss\subset \wt{\mfg}_\reg \subset \wt\mfg$ be their preimages.

		Let $\mCN_\reg:=\mfg_\reg\cap \mCN$ be the open subscheme of $\mCN$. Its closed points are regular nilpotent elements.

		Let $\mft_\reg^\abs\subset \mft^\abs$ be the open subscheme whose closed points are regular elements\footnote{This means the realizations in any/all Cartan subalgebras are regular. Equivalently, this means $W^\abs$ acts freely at these points.}.
	\end{notn}

	We have the following basic results. See e.g. \cite[Section 3.1]{CG} for a proof.

	\begin{prop}
		\label{prop-Groth-alt}
		Consider the map $\mfg \to \mfg/\!/G \to \mft^\abs/\!/W^\abs$ given by the abstract Chevalley isomorphism (see Appendix \ref{app-abs}). We have:
		\begin{itemize}
			\item[(1)]
				The following diagram commutes:
				\[
					\xymatrix{
						\wt\mfg \ar[r] \ar[d] & \mft^\abs \ar[d] \\
						\mfg \ar[r] & \mft^\abs/\!/W^\abs
					}
				\]
			\item[(2)]
				When restricted to the regular locus $\wt\mfg_\reg$, the above diagram is Cartesian. In other words, the following diagram is Cartesian:
				\[
					\xymatrix{
						\wt\mfg_\reg \ar[r] \ar[d] & \mft^\abs \ar[d] \\
						\mfg_\reg \ar[r] & \mft^\abs/\!/W^\abs.
					}
				\]
			\item[(3)]
				When restricted to the regular semisimple locus $\wt\mfg_\rss$, the following diagram is Cartesian, and the Vertical maps are finite étale covers with Galois group $W^\abs$:
				\[
					\xymatrix{
						\wt\mfg_\rss \ar[r] \ar[d] & \mft_\reg^\abs \ar[d] \\
						\mfg_\rss \ar[r] & \mft_\reg^\abs/W^\abs.
					}
				\]
		\end{itemize}

	\end{prop}

	\begin{warn}
		The map $\wt\mfg \to \mft^\abs$ does \emph{not} send regular elements to regular elements. Indeed, it sends $\mCN_\reg$ to $0$.
	\end{warn}


	
	\begin{prop}
		The scheme $\mCN$ is normal.
	\end{prop}

	\proof[Sketch]
		We have seen $\mCN$ is Cohen--Macaulay (Remark \ref{rem-Nilp-CM}). Hence by Serre's criterion, we only need to show $\mCN$ is regular in codimension $1$. 

		By Lemma \ref{lem-Groth-alter-Cartan}, the map $\wt{\mfg} \to \mft^\abs$ is smooth, hence so is $\wt{\mfg}_\reg \to \mft^\abs$. Recall $\mft^\abs\to \mft^\abs/\!/W$ is faithfully flat ([Lecture 6, Proposition 1.1 and Corollary 1.5]). Hence by Proposition \ref{prop-Groth-alt}(2), the map $\mfg_\reg \to \mft^\abs/\!/W^\abs$ is smooth (by flat descent of smooth maps). By definition, we have a Cartesian diagram
		\[
			\xymatrix{
				\mCN_\reg \ar[r] \ar[d] & \mfg_\reg \ar[d] \\
				0 \ar[r] & \mfg/\!/G \simeq \mft^\abs/\!/W^\abs.
			}
		\]
		Hence $\mCN_\reg$ is smooth.

		It remains to show the closed subset $\mCN-\mCN_\reg$ of $\mCN$ is of codimension $\ge 2$. Since $\mCN$ is irreducible (Corollary \ref{cor-Nilp-irre}), $\mCN-\mCN_\reg$ is of codimension $\ge 1$. We need to use the following two basic facts:
		\begin{itemize}
			\item[(i)]
				The adjoint action of $G$ on $\mCN$ has only finitely many orbits\footnote{This can be viewed as a generalization of the theory of Jordan blocks.} (see \cite[Proposition 3.2.9]{CG});
			\item[(ii)]
				Each $G$-orbit on $\mfg$ has a symplectic structure (see \cite[Proposition 1.1.5]{CG}).
		\end{itemize}
		By (ii), each $G$-orbit has an even dimension. Hence each $G$-orbit in $\mCN-\mCN_\reg$ has even codimension. By (i), $\mCN-\mCN_\reg$ has codimension $\ge 2$ as desired.

	\qed

	\begin{cor}
		The map $\mfp:\wt\mCN \to \mCN$ is a resolution of singularities, i.e., it is birational proper and surjective.
	\end{cor}

	\proof
		We have already proved $\mfp$ is proper and surjective (Lemma \ref{lem-Springer-proper-surj}). It remains to show $\mfp$ is birational. We claim its restriction on $\mCN_\reg\subset \mCN$ is an isomorphism. Since $\mCN$ is reduced, we only need to show any closed point of $\mCN_\reg$ has a unique preimage in $\wt\mCN_\reg$. Now the claim follows from Proposition \ref{prop-Groth-alt}(2) because $0\in \mft^\abs/\!/W^\abs$ has a unique preimage in $\mft^\abs$.

	\qed

	\begin{rem}
		In fact, $\mfp:\wt\mCN \to \mCN$ is a \emph{semismall} resolution, i.e., $\dim(\wt\mCN\mt_\mCN\wt\mCN) = \dim(\mCN)$. This fact is crucial in the Springer theory. The fiber product $\wt\mCN\mt_\mCN\wt\mCN$ is known as the \textbf{Steinberg variety}, which also plays a central role in geometric representation theory. For more information, see \cite{CG}.
	\end{rem}

	\proof[Proof of Theorem \ref{thm-Kostant-Springer}]

		Follows by applying Zariski's main theorem to the projection $\mfp: \wt\mCN \to \mCN$. Direct proof: $\mfp_* \mCO_{\wt\mCN}$ is coherent because $\mfp$ is proper. It is generically of rank $1$ because $\mfp$ is birational. Both the source and target of $\mCO_\mCN\to \mfp_* \mCO_{\wt\mCN}$ are sheaves of integral domains, hence they have isomorphic sheaves of fractional fields. Then we win because $\mCO_\mCN$ is integrally closed.

	\qed


	\appendix

	\section{Abstract Cartan group and abstract Weyl group}
	\label{app-abs}

	\begin{constr}
		Let $B_x$ and $B_y$ be two Borel subgroups of $G$. Let $T_x:=B_x/[B_x,B_x]$ and $T_y:=B_y/[B_y,B_y]$ be their abelianizations. Recall there exists $g\in G(k)$ such that $\Ad_g$ induces an isomorphism $\Ad_g: B_x \xrightarrow{\simeq} B_y$. Hence we obtain an isomorphism between the abelianizations $\ol{\Ad}_g: T_x \xrightarrow{\simeq} T_y$.
		The isomorphism $\ol{\Ad}_g$ does not depend on the choice of $g$ because any other choice $g'$ satisfies $g'\in g B_x(k)$ and the adjoint action of $B_x$ on $T_x$ is trivial. For this reason, we write the above isomorphism as
		\[
			\phi_{x,y}: T_x \xrightarrow{\sim} T_y.
		\]

		It is easy to check $\phi_{x,x}=\Id$ and $\phi_{y,z}\circ \phi_{x,y}=\phi_{x,z}$. Hence there exists an algebraic group $T^\abs$, equipped with isomorphisms
		\[
			r_x: T^\abs \xrightarrow{\sim} T_x,
		\]
		such that $r_y = \phi_{x,y}\circ r_x$. The data $(T^\abs, r_x)$ are \emph{unique up to an unique isomorphism}\footnote{This means for $(T^\abs, r_x)$ and $((T^\abs)', r_x')$, there is a unique isomorphism $\alpha:T^\abs \xrightarrow{\sim} (T^\abs)'$ such that $r_x= r_x'\circ \alpha$.}.

		We call $T^\abs$ the \textbf{abstract Cantan group} for $G$, and call $r_x$ the \textbf{realizing isomorphisms}.

		Similarly, the Lie algebra of $T^\abs$ is denoted by $\mfg^\abs$ and is called the \textbf{abstract Cantan algebra} for $\mfg$.
	\end{constr}

	\begin{warn}
		The algebraic group $T^\abs$ is \emph{not} a subgroup of $G$, at least not in a canonical way.
	\end{warn}

	\begin{rem}
		A Cartan subgroup $T_1\inj G$ of $G$ does \emph{not} give a realizing isomorphism $T^\abs \to T_1$, at least not in a canonical way. Instead, if we further choose a Borel subgroup $B_x$ that contains $T_1$, i.e., if we have a \textbf{Borel pair} $(B_x,T_1)$, then there is a realizing isomorphism $T^\abs \to T_1$ defined to be the composition
		\[
			r_{(B_x,T_1)}:T^\abs \xrightarrow{r_x} T_x \xleftarrow{\sim} T_1,
		\]
		where the second isomorphism is given by $T_1 \inj B_x \surj T_x$.
	\end{rem}

	\begin{warn}
		One cannot define the abstract Borel group for $G$.
	\end{warn}

	\begin{constr}
		Let $(B_x,T_1)$ and $(B_y,T_2)$ be two Borel pairs. Recall there is a unique element $g\in G(k)$ such that $\Ad_g: B_x\xrightarrow{\sim} B_y$ and $\Ad_g: T_1 \xrightarrow{\sim} T_2$. Hence we obtain an isomorphism between the normalizers $\Ad_g: N_G(T_1) \xrightarrow{\sim} N_G(T_2)$ and therefore an isomorphism between the corresponding Weyl groups. We denote this isomorphism by
		\[
			\varphi_{(B_x,T_1),(B_y,T_2)}: W_{T_1} \to W_{T_2}.
		\]

		It is easy to check $\varphi_{(B_x,T_1),(B_x,T_1)}=\Id$ and $ \varphi_{(B_y,T_2),(B_z,T_3)}\circ \varphi_{(B_x,T_1),(B_y,T_2)} = \varphi_{(B_x,T_1),(B_z,T_3)} $. Hence there exists a group $W^\abs$ equipped with isomorphisms
		\[
			r_{(B_x,T_1)}: W^\abs \xrightarrow{\sim} W_{T_1}
		\]
		such that $r_{(B_y,T_2)} = \varphi_{(B_x,T_1),(B_y,T_2)}\circ r_{(B_x,T_1)}$. The data $(W^\abs, r_{(B_x,T_1)})$ are unique up to an unique isomorphism.

		We call $W^\abs$ the \textbf{abstract Weyl group} for $G$, and call $r_{(B_x,T_1)}$ the \textbf{realizing isomorphisms}.

	\end{constr}

	\begin{warn}
		The isomorphism $\varphi_{(B_x,T_1),(B_y,T_2)}$ depends on $B_x$ and $B_y$.
	\end{warn}

	\begin{warn}
		The group $W^\abs$ is \emph{not} a subgroup of $G$, at least not in a canonical way.
	\end{warn}

	\begin{rem}
		One can also define the abstract Weyl group by providing a group structure on $|G\backslash (X\mt X)|$. This construction was introduced by Deligne--Lusztig when developing the theory named by them.

	\end{rem}

	\begin{constr}
		Let $(B_x,T_1)$ be any Borel pair. The action of $W_{T_1}$ on $T_1$ defines an action of $W^\abs$ on $T^\abs$ via the realizing isomorphisms $r_{(B_x,T_1)}: T^\abs \xrightarrow{\sim} T_1$ and $r_{(B_x,T_1)}: W^\abs \xrightarrow{\sim} W_{T_1}$. Unwinding the definitions, one can show this action does not depend on the choice of the Borel pair. Hence we obtain a \emph{canonical} action of $W^\abs$ on $T^\abs$, which is called \textbf{the (abstract) action} of $W^\abs$ on $T^\abs$.
	\end{constr}

	\begin{constr}
		Recall for any Borel pair $(B,T)$, we have the Chevalley isomorphism $\mft/\!/W \xrightarrow{\sim} g/\!/G$ characterized by the following commutative diagram (see [Lecture 5, (4.2)])
		\[
			\xymatrix{
				\mfb \ar[r] \ar[d] & \mfg/\!/G \\
				\mft \ar[r] & \mft/\!/W. \ar[u]_-\simeq
			}
		\]
		Via the realizing isomorphism $r_{(B,T)}: \mft^\abs/\!/W^\abs \xrightarrow{\sim} \mft/\!/W$, we obtain an isomorphism 
		\[
			\mft^\abs/\!/W^\abs \xrightarrow{\sim} \mfg/\!/G
		\]
		which can be shown to do not depend on the choice of the Borel pair. We call this isomorphism the \textbf{abstract Chevalley isomorphism}.
	\end{constr}



	
	
	

\begin{thebibliography}{Yau}

	\bibitem[CG]{CG} Chriss, Neil, and Victor Ginzburg. Representation theory and complex geometry. Vol. 42. Boston: Birkhäuser, 1997.
	

	\bibitem[G]{G} Gaitsgory, Dennis. Course Notes for Geometric Representation Theory, 2005, available at \url{https://people.mpim-bonn.mpg.de/gaitsgde/267y/catO.pdf}.

\end{thebibliography}


\end{document} 


